---
title: "HW7"
author: "Xinrui Hu"
date: "2023-11-10"
output: word_document
---

```{r}
library(rio)
marij <-import(file = "Ch06_NELS88_Marijuanause.dta")
marij$MJ[marij$MJ > 3] <- NA
library(VGAM)
```

```{r}
# Ex2
fit_zach<-vglm(factor(MJ)~ ZACH,
family = multinomial(refLevel = 1), data = marij)
summary(fit_zach)
```

```{r}
fit_zrr<-exp(coefficients(fit_zach))
fit_zrr
#The relative risk ratio (RRR) of using MJ on 3-19 occasions vs never used MJ decreases by a factor of(1-0.818) * 100% = 18.2% when ZACH increases 1 SD. 
# The odds of using of MJ on 3-19 occasions among students with averaged achievement is 0.08
#The relative risk ratio (RRR) of using MJ on 20 or more occasions vs never used MJ decreases by a factor of1-0.747=25.3% when ZACH increases 1 SD.
#The odds of using of MJ on 20 or more occasions among students with averaged achievement is 0.05
```
Ex3

```{r}
# Ex4
max(marij$ZACH, na.rm = TRUE)
min(marij$ZACH, na.rm = TRUE)
which.max(marij$ZACH)
which.min(marij$ZACH)
```

```{r}
predictedall <- fitted(fit_zach, na.rm = FALSE)
predictedall["193",]
#predictedall["21477",]
#PI0 is 0.72
#PI1 is 0.13
#PI2 is 0.087
#PI3 is 0.066
#Min Sum is 0.72 + 0.13 + 0.087 + 0.066 = 1
#Max Sum is 0.88 + 0.057 + 0.044 + 0.022 = 1
```

```{r}
# Exercise 5
library(haven)
Spank <- read_dta("Lecture 8_Spanking_Gender.dta")
fit_spank <- vglm(spanking ~ gender + educate + polviews + zage, family = cumulative(parallel = TRUE), data = Spank)
summary(fit_spank)
```
After controlling for the effects of education, zscore of age and political views, the
expected odds of strongly disagreeing are exp(-0.25) = 0.775 times
among females versus among males.

After controlling for the effects of gender, zscore of age and political views, each one-unit
increase in education is associated with a {[1 - exp(-.11)]* 100%} = 11% decrease
in the odds of strongly disagreeing.

After controlling for the effects of gender, zscore of age and
education, each one-unit increase in political view
is associated with a {[exp(.22)-1]* 100%} = 24%
increase in the odds of strongly disagreeing

After controlling for the effects of gender, political views and
education, each one-unit increase in z score of age
is associated with a {[exp(.07)-1]* 100%} = 7%
increase in the odds of strongly disagreeing
```{r}
# 8.2
# a)
tab <- matrix(c(833, 125, 2, 160), ncol=2, byrow=T)
tab
mcnemar.test(tab, correct=F)
# Since p-value < 0.0001, the evidence is strong enough to reject the null hypothesis and the population proportion answering “yes” weren’t identical for heaven and hell.
```
```{r}
#b)
total = 162+125+833
diff = 285/(total)-162/(total)
SE = sqrt(127-(123^2)/(total))/(total)
diff + c(-1, 1)*1.96*SE
#We are 95% confidence that the probability of “yes” is at least 0.0912 and at most 0.128 for believe in heaven than in hell.
```
```{r}
#c)
abs(1-exp(log(((125+833)/(total-125-833))/((833+2)/(total-833-2)))))
125/2
# The odds of people belive in hevean is 101.8% higher than the belive in hell
#The subject odds of people belief in heaven is 62.5 times higher than the subject odds of belief in hell.
# Marginal models offer an average interpretation across the population, while subject-specific models provide an interpretation conditional on the random effects 
```

